Group Signatures Meet Anonymous Credentials: Or, How to Build Privacy Preserving Identities with Accountability

### 2022-02-15 prism-backend/research/gsac/uas-model.tex
Refined model of UAS. Working on properties.

\section{Model for \GSAC Schemes}
\label{sec:model}

\todo{Big doubt in this model. Currently, \Open (and \OPEN) return the public
  key and credential of the signer along with the correctness proof. This is
  as opposed to conventional practice, in which just an index and the proof
  are returned. It is also intentional, as in the future I'd
  like to extend \Open to a sort of functional version, where instead we return
  a function of the user (secret or public) key, the credential, and the
  signed message. However, I am not sure yet if this will be too strong to be
  realized. If it is possible, though, our model would be stronger in this
  sense. And also, I like this better, as by outputing an index, but then
  requiring \Judge to use the public key (as in \cite{bsz05}) you impose the
  need to create ad hoc protocols so that the \Judge learns the actual public
  key, which in the real world can lead to very bad practices... (Such as
  integrating the Judge inside the Opener, and not making the correctness
  proof public -- hence, full trust in the Opener+Judge; or implementations
  of \Open that directly output the public key instead of the index for
  convenience, but in that case, deviating from the model and thus without
  security guarantees.)}

While group signatures and anonymous credentials share quite common syntax and
security properties, which are even frequently formalised in a similar way,
there are aspects that need to be considered with care.

The first choice that has to be made is whether to make sign/show and
verification a non-interactive or an interactive process -- in group signatures,
we are in the former case, whereas in anonymous credentials, we are in the
latter. While both options are of course valid, this has direct impact in the
modelling. We choose the non-interactive approach, and then prove how to
generically translate it into an interactive process \todo{do we?}. The reason
is that this results in a more flexible building block, suitable for a wider set
of scenarios. And, syntactically, it seems more natural to combine two
non-interactive processes into an interactive one, than the other way around.

We also need to take into account that, in group signature schemes, users only
get one membership credential -- typically bound to their personal secret
key --, which they then use to create as many group signatures as they want. On
the other hand, in anonymous credentials, users can get as many credentials as
they desire, which can (typically) then be showed also arbitrarily many times.
In this regard, we follow a somehow intermediate approach: users own a single
personal secret key, which they can use to get as many credentials as they wish.
Subsequently, they can use any of those credentials to create \GSAC signatures.
This has, as expected, some nuanced impact in the modelling. For instance, in
group signatures, tracing and non-frameability are frequently dependent on the
open function, which (among other things) returns an index pointing to the join
transcript associated with the user who produced the signature given by the
adversary as a response to the game's challenge. In our approach, though, each
user may have many such transcripts -- one per credential that the user
obtains. However, note that what actually provides meaning to these notions
is not the index; but the associated user key pair, which is moreover necessary
to prove that the opening has been performed honestly. Thus, with some
restructuring of the corresponding games, we can still capture the main
meaning of these traceability and non-frameability notions from group
signatures. This also impacts anonymity, since in group signatures, this is
modelled by challenging the adversary to guess which user (which implicitly
translates to ``which user credential'') out of two challenge users was
leveraged to produce the challenge signature. But again, in our case, users may
have multiple credentials. However, it seems sufficient to adopt a similar
anonymity flavour to that of anonymous credentials, where the adversary has to
distinguish between presentations (signatures) produced from two arbitrary
credentials, independently on whether these credentials were obtained by the
same user, or not.

Getting closer to the anonymity definition of anonymous credentials also has the
side effect to introduce the notions of attributes for free into the domain of
group signatures -- which, up to now, had considered membership credentials
\emph{without} attributes. Yet another side effect of introducing multiple
attributes (if we think from a group signature perspective) is that, in order
to avoid trivial wins, the anonymity game has to enforce that both challenge
credentials reveal the same predicate on their attributes.
\commentwho{Jesus}{Again, taking into account ABGS, this is no longer true.
  Cross-check and re-write.}

In the model described below, we follow closely the definitions from
\cite{bsz05} for group signatures, and \cite{fhs19} for anonymous credentials
(which are in turn influenced by \cite{bsz05}).

@@

\begin{description}
\item[$\parm \gets \Setup(\secpar)$.] Given a security parameter \secpar,
  returns a global system parameter variable \parm.
\item[$(\ipk,\isk) \gets \IKeyGen(\parm)$.] Given global system parameters
  \parm, returns the issuer's key pair. Hereafter, we assume that the public
  part \ipk is added to the group public key \gpk.
\item[$(\opk,\osk) \gets \OKeyGen(\parm)$.] Given global system parameters
  \parm, returns an opener's key pair. Hereafter, we assume that the public part
  \opk is added to the group public key \gpk.
\item[$(\upk,\usk) \gets \UKeyGen(\parm)$.] Given global system parameters
  \parm, returns a user's key pair.
\item[$\langle \cred/\bot,\utrans/\bot \rangle \gets
  \langle \Obtain(\gpk,\usk,\Attrs),\Issue(\gpk,\isk,\upk,\Attrs) \rangle$.]
  This interactive protocol lets a user with key pair (\upk,\usk) running the
  \Obtain process, obtain a credential \cred with attributes \Attrs, by
  communicating with an issuer, with private key \isk, running the \Issue
  counterpart. The user outputs the produced credential \cred, while the issuer
  outputs the protocol transcript \utrans for the produced credential.
\item[$\sig \gets \Sign(\gpk,\usk,\cred,\DAttrs,\msg)$.] The user with
  with secret key \usk, who obtained a credential \cred, produces a signature
  \sig over message \msg, revealing subset of attributes \DAttrs in \cred.
\item[$1/0 \gets \Verify(\gpk,\sig,\DAttrs,\msg)$.] Checks whether \sig is a
  valid signature revealing the attribute set \DAttrs of its signing credential,
  over message \msg.
\item[$(\upk,\cred,\oproof)/\bot \gets
  \Open(\gpk,\osk,\utrans,\sig,\DAttrs,\msg)$.]
  Given a signature \sig over message \msg, produced by a credential with
  attributes containing the set \DAttrs, associated to transcript \utrans,
  returns the user public key that was used to obtain the credential used for
  producing \sig, the credential, and a proof of opening correctness \oproof; or
  $\bot$ if the opening process fails.  
\item[$1/0 \gets \Judge(\gpk,\upk,\cred,\oproof,\sig,\DAttrs,\msg)$.] Checks if
  \oproof is a valid opening correctness proof stating that the user with public
  key \upk generated the signature \sig over message \msg, by using credential
  \cred with attributes containing the set \DAttrs. 
\end{description}

@@

\paragraph{Global Variables.} %
All the honestly generated users (i.e., all honestly generated user key pairs,
since we assume a one-to-one relationship between user and user key pair), as
well as all honestly generated credentials, are assigned identifiers. Wex
typically write as \uid for users' identifiers, and \cid for credentials'. The
adversary can
refer to any individual user or credential using the corresponding identifier --
even though he may not know the actual contents of the key pairs or credentials.
In games that involve challenge user or credential identifiers, we use \cuid and
\ccid to refer to these challenge user/credential.
%
The games, usually through the oracles, keep track of the users (through
set \USK) and credentials (through table \CRED) that are created as a result of
oracle calls by the adversary, as well as the attributes that each credential
is associated (through table \ATTR), and the user to which each credential has
been issued to (through table \OWNR). We use the user and credential identifiers
to reference specific user key pairs, credentials, and credential attributes.
For instance, $\USK[\uid]$ refers to the user key pair corresponding to user
with identifier \uid, $\PUBUSK[\uid]$ refers to the public key of that key pair,
and $\PRVUSK[\uid]$ to the private key; credential $\CRED[\cid]$ refers to the
credential data associated to the credential with identifier \cid; $\ATTR[\cid]$
is the set of attributes that was assigned to the credential with identifier
\cid; and $\OWNR[\cid]$ is set to the \uid corresponding to the user that
credential \cid was issued to.
%
Additionally, the games keep track of all honest and corrupt users that have
been generated, through sets \HU and \CU, respectively. They also keep track of
the signatures that have been honestly produced, through the \SIG table. Since
signatures are produced through credentials, the \SIG table is indexed with
{\cid}s. For the anonymity game, we also need to keep track of the challenge
signatures that the adversary has obtained, in order to prevent trivial wins
by allowing any of them to be opened. 
%
All global variables are initially set by the games to empty values (denoted
with $\bot$), and all tables/sets are initialized as empty tables/sets, denoted
with $\emptyset$. Also, for readability, we abuse the syntax as follows: we
write $\CRED[\uid]$ to mean $\CRED[\cid]$ for all $\cid$ such that
$\OWNR[\cid] = \uid$; \todo{Something else?}

\paragraph{Oracles.} %
Oracles are the interface of the adversary with the corresponding games. In
other words: through these oracles, the game enviroment exposes to the adversary
functionality that could otherwise be executed only by honest parties with
private knowledge -- knowledge that would make the adversary capable of
trivially breaking the security properties formalized in the experiments.
In the game-based definitions of our \GSAC model, we leverage the following
oracles, which are formally defined in \figref{fig:oracles}

\begin{description}
\item[$\HUGEN$.] Adds a new honest user to the game, by honestly generating
  the user's key pair.
\item[$\CUGEN$.] If the specified user identifier is not already in the game,
  sets the user public key to a value given by the adversary. If it already
  exists, and is associated to an honest user, then it reveals to the adversary
  the user's secret key.
\item[$\OBTISS$.] Lets the adversary add a new honestly generated credential to
  the game, on behalf of an honest user.
\item[$\OBTAIN$.] Enables the adversary to play the role of a dishonest issuer
  in games that support it, by letting it interact with honest users who want to
  receive credentials.
\item[$\ISSUE$.] Allows the adversary to play the role of dishonest users,
  requesting an honest issuer to produce credentials fo them.
\item[$\SIGN$.] Lets the adversary get signatures from credentials belonging
  to honest users.
\item[$\OPEN$.] Given an honestly produced signature, lets the adversary learn
  which credential was used to produce it.
\item[$\CHALb$.] Upon receiving two challenge credentials, a common intersecting
  set of attributes, and a message, returns a signature produced by one of these
  two credentials, defined by the bit $b$, which is established in the anonymity
  game.
\end{description}

@@

\paragraph{Non-frameability.} %
Non-frameability variants are a core unforgeability-type property in group
signatures. However, no
similar property is modelled for anonymous credentials (\todo{see the discussion
  in \secref{sec:introduction} for further detail}). It is a quite strong
property, as it must be ensured even in the presence of dishonest issuer and
opener. Intuitively, it prevents the adversary from creating a signature that
frames an honest user. Depending on the inspection capabilities of the scheme,
this framing could be done in different ways; i.e., by convincing third parties
that signatures by different (possibly corrupt) users are linked, or directly
by having open proofs output the identity of a user who did not create the
signature being opened.
%
In \GSAC schemes, in order for a user to be framed, the adversary first needs to
create a $(\sig,\DAttrs,\msg,\upk,\cred,\oproof)$ such that the signature,
attributes, and message are accepted by \Verify, and which, along with the user
public key, credential, and open correctness proof, are accepted by \Judge.
Then, the adversary wins the game if the owner of the credential is honest, but
the signature was not produced via the \SIGN oracle, or if the public key output
by the adversary does not match the public key associated to the credential
owner.
%
Note that, as in anonymity, this notion of non-frameability is even stronger
to that of group signatures, since our \Open function returns both the public
key and the credential that were used to create the signature.

@@

\subsection{Tranformation to Interactive Presentations}
\label{ssec:interactivetransform}

>>>>

\section{Model for \UAS Schemes}
\label{sec:model-uas}

Universal Anonymous Signatures.

@@

In the following, we assume a setting with multiple groups. For simplicity,
we assume that each has its own issuer and inspector.

\todo{Are generalizations where different groups can share issuer or inspector
  straight forward?} \\
\todo{The notation for multisets is not strictly correct. Still, leaving it as
  is for now.} \\
\commentwho{Jesus}{I'm trying to define \Sign, \Verify, \Inspect and \Judge such
  that multiple credentials can be used to produce a single signature, proving
  claims over arbitrary subsets of all the credentials owned by a user. \Verify
  and, consequently, \Inspect and \Judge need to receive all the \gpk's.
  \Inspect only requires one inspector's secret key, though, and will thus only
  produce that inspector's
  function of the signature and message (this is to avoid complex interactive
  protocols between many inspectors). This, however, creates a quite annoying
  syntax. It may just be easier to restrict signatures to claims over the
  attributes contained in \emph{one} credential.}

\begin{description}
\item[$\parm \gets \Setup(\secpar)$.] Given a security parameter \secpar,
  returns a global system parameter variable \parm.
\item[$(\ipk,\isk) \gets \IKeyGen(\parm,\LangIss)$.] Given global system
  parameters \parm, and the language \LangIss defining the NP relation for
  issuance, an issuer runs this function to generate its issuing key pair.
  Hereafter, we assume that the public part \ipk is added to the group public
  key \gpk.
\item[$(\opk,\osk) \gets \OKeyGen(\parm,\LangIns)$.] Given global system
  parameters \parm, and the language \LangIns defining the NP relaton for
  inspection, an inspector runs this function to generate its inspecting key
  pair. It returns the inspector's key pair. Hereafter, we assume that the
  public part \opk is added to the group public key \gpk.
\item[$(\upk,\usk) \gets \UKeyGen(\parm)$.] Given global system parameters
  \parm, returns a user's key pair.
\item[$\langle \cred/\bot,\utrans/\bot \rangle \gets
  \langle
  \Obtain(\usk,\UAttrs,\DAttrs,\ldblbrace (gpk_i,\cred_i,\UAttrs_i,\DAttrs_i)\rdblbrace_{i \in \Issuers}),
  \Issue(\isk,\DAttrs,\upk,\ldblbrace (\gpk_i,\DAttrs_i)\rdblbrace_{i \in \Issuers})
  \rangle$.] %
  This interactive protocol lets a user with key pair (\upk,\usk) running the
  \Obtain process, obtain a credential \cred from an issuer in the system, on
  attribute set $\UAttrs \cup \DAttrs$, where the attributes in $\UAttrs$ are
  kept hidden from the issuer, while those in \DAttrs are revealed. In order
  to get the credential, the user may leverage previously obtained credentials
  $\cred_i$ from a multiset of other issuers (including the one to whom a new
  credential is being requested) with public keys $\gpk_i$. Moreover, each
  credential $\cred_i$ can be used to prove claims over undisclosed attributes
  $\UAttrs_i$ that the user wishes to include in the new credential. The user
  outputs the produced credential \cred, while the issuer outputs the protocol
  transcript \utrans for the produced credential.
\item[$\sig \gets \Sign(\usk,\ldblbrace (\gpk_i,\cred_i,\UAttrs_i,\DAttrs_i)
  \rdblbrace_{i \in \Issuers},\msg)$.] %
  The user with with secret key \usk, who obtained credentials $\cred_i$ from a
  multiset of issuers in the system, produces a signature \sig over message
  \msg. The produced signature proves arbitrary claims over the set of
  disclosed attributes $\bigcup_i \DAttrs_i$ and undisclosed attributes
  $\bigcup_i \UAttrs_i$.
\item[$1/0 \gets
  \Verify(\sig,\msg,\ldblbrace (\gpk_i,\DAttrs_i) \rdblbrace_{i \in \Issuers})$.]
  Checks whether \sig is a valid signature, over message \msg, satisfying
  arbitrary claims over disclosed attributes $\bigcup_i \DAttrs_i$, as well as
  potentially other undisclosed attributes endorsed by issuers in the systsem
  with public keys $\bigcup_i \gpk_i$.
\item[$(\y,\iproof)/\bot \gets
  \Inspect(\osk,\utrans,\sig,\msg,
  \ldblbrace (\gpk_i,\DAttrs_i) \rdblbrace_{i \in \Issuers})$.] %
  \commentwho{Jesus}{I do not like this definition. It's complicated. This needs
    to be rephrased from scratch.}
  Executed by the inspector with opening key \osk. Receives a signature \sig
  over message \msg, produced by a set of credentials satisfying claims on
  disclosed attributes $\bigcup_i \DAttrs_i$, as well as potentially undisclosed
  attributes, all endorsed by a multiset of issuers with public keys $\gpk_i$. 
  If \utrans is a valid transcript of the execution of a
  $\langle\Obtain,\Issue\rangle$ interactive protocol where the signer obtained
  the credential associated to the inspector's group, then the function outputs
  a value $\y$ derived from the signer's credential and public key, and the
  signed message, as well as a proof of correct inspection.
\item[$1/0 \gets \Judge(\y,\iproof,\sig,\msg,
  \ldblbrace (\gpk_i,\DAttrs_i)\rdblbrace_{i \in \Issuers})$.] %
  Checks if \iproof is a valid inspection correctness proof for the value \y,
  obtained by applying \Inspect to the the pair signature \sig over \msg. 
\end{description}

@@

\paragraph{Global Variables.} %
All the honestly generated users (i.e., all honestly generated user key pairs,
since we assume a one-to-one relationship between user and user key pair), as
well as all honestly generated credentials, are assigned identifiers. Week's
typically write as \uid for users' identifiers, and \cid for credentials'. The
adversary can
refer to any individual user or credential using the corresponding identifier --
even though he may not know the actual contents of the key pairs or credentials.
In games that involve challenge user or credential identifiers, we use \cuid and
\ccid to refer to these challenge user/credential.
%
The games, usually through the oracles, keep track of the users (through
set \USK) and credentials (through table \CRED) that are created as a result of
oracle calls by the adversary, as well as the attributes that each credential
is associated (through table \ATTR), and the user to which each credential has
been issued to (through table \OWNR). We use the user and credential identifiers
to reference specific user key pairs, credentials, and credential attributes.
For instance, $\USK[\uid]$ refers to the user key pair corresponding to user
with identifier \uid, $\PUBUSK[\uid]$ refers to the public key of that key pair,
and $\PRVUSK[\uid]$ to the private key; credential $\CRED[\cid]$ refers to the
credential data associated to the credential with identifier \cid; $\ATTR[\cid]$
is the set of attributes that was assigned to the credential with identifier
\cid; and $\OWNR[\cid]$ is set to the \uid corresponding to the user that
credential \cid was issued to.
%
Additionally, the games keep track of all honest and corrupt users that have
been generated, through sets \HU and \CU, respectively. They also keep track of
the signatures that have been honestly produced, through the \SIG table. Since
signatures are produced through credentials, the \SIG table is indexed with
{\cid}s. For the anonymity game, we also need to keep track of the challenge
signatures that the adversary has obtained, in order to prevent trivial wins
by allowing any of them to be opened. 
%
All global variables are initially set by the games to empty values (denoted
with $\bot$), and all tables/sets are initialized as empty tables/sets, denoted
with $\emptyset$. Also, for readability, we abuse the syntax as follows: we
write $\CRED[\uid]$ to mean $\CRED[\cid]$ for all $\cid$ such that
$\OWNR[\cid] = \uid$; \todo{Something else?}

\paragraph{Oracles.} %
Oracles are the interface of the adversary with the corresponding games. In
other words: through these oracles, the game environment exposes to the adversary
functionality that could otherwise be executed only by honest parties with
private knowledge -- knowledge that would make the adversary capable of
trivially breaking the security properties formalized in the experiments.
In the game-based definitions of our \GSAC model, we leverage the following
oracles, which are formally defined in \figref{fig:oracles}

\begin{description}
\item[$\HUGEN$.] Adds a new honest user to the game, by honestly generating
  the user's key pair.
\item[$\CUGEN$.] If the specified user identifier is not already in the game,
  sets the user public key to a value given by the adversary. If it already
  exists, and is associated to an honest user, then it reveals to the adversary
  the user's secret key.
\item[$\OBTISS$.] Lets the adversary add a new honestly generated credential to
  the game, on behalf of an honest user.
\item[$\OBTAIN$.] Enables the adversary to play the role of a dishonest issuer
  in games that support it, by letting it interact with honest users who want to
  receive credentials.
\item[$\ISSUE$.] Allows the adversary to play the role of dishonest users,
  requesting an honest issuer to produce credentials for them.
\item[$\SIGN$.] Lets the adversary get signatures from credentials belonging
  to honest users.
\item[$\OPEN$.] Given an honestly produced signature, lets the adversary learn
  which credential was used to produce it.
\item[$\CHALb$.] Upon receiving two challenge credentials, a common intersecting
  set of attributes, and a message, returns a signature produced by one of these
  two credentials, defined by the bit $b$, which is established in the anonymity
  game.
\end{description}

@@

\paragraph{Non-frameability.} %
Non-frameability variants are a core unforgeability-type property in group
signatures. However, no
similar property is modeled for anonymous credentials (\todo{see the discussion
  in \secref{sec:introduction} for further detail}). It is a quite strong
property, as it must be ensured even in the presence of dishonest issuer and
opener. Intuitively, it prevents the adversary from creating a signature that
frames an honest user. Depending on the inspection capabilities of the scheme,
this framing could be done in different ways; i.e., by convincing third parties
that signatures by different (possibly corrupt) users are linked, or directly
by having open proofs output the identity of a user who did not create the
signature being opened.
%
In \GSAC schemes, in order for a user to be framed, the adversary first needs to
create a $(\sig,\DAttrs,\msg,\upk,\cred,\oproof)$ such that the signature,
attributes, and message are accepted by \Verify, and which, along with the user
public key, credential, and open correctness proof, are accepted by \Judge.
Then, the adversary wins the game if the owner of the credential is honest, but
the signature was not produced via the \SIGN oracle, or if the public key output
by the adversary does not match the public key associated to the credential
owner.
%
Note that, as in anonymity, this notion of non-frameability is even stronger
to that of group signatures, since our \Open function returns both the public
key and the credential that were used to create the signature.

@@

\subsection{Transformation to Interactive Presentations}
\label{ssec:interactivetransform}
